---
title: "Mountain Yellow-legged Frog Population Assesment"
author: "Conner Smith"
date: "1/13/2022"
output: 
  html_document: 
    theme: flatly
---

## **Overview**

This task explores the occurrence of endangered Mountain Yellow-legged frogs in lakes throughout the Sierra Nevada region. Counts are drawn from amphibian abundance data recorded by the Sierra Lakes Inventory Project which ran from 1995-2002 and has supported research and management of Sierra Nevada aquatic ecosystems and their terrestrial interfaces. 

```{r setup, include=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(cowplot)
library(patchwork)
library(lubridate)

```

## **Analysis**

This analysis filters the larger dataset to gather the counts for Moutain Yellow-legged frogs by life stage (Adult, Sub-adult, or Tadpole) across the region as a whole. It also includes the total counts (irrespective of age but excluding tadpoles) for the five lakes with the highest recorded numbers of this species and presents the results ina  combined visualization (**Figure 1**). 

```{r}
# Read in xlsx file with 'read_xlsx

frogs <- readxl::read_xlsx(here("data", "sierra_amphibians.xlsx"))
```


```{r}
# Filter for Mountain Yellow-legged Frogs (Species code: RAMU), excluding egg life stage (code: EggMass) and group by year and lifestage. 

frogs_clean <- frogs %>% 
  filter(amphibian_species == 'RAMU',
         amphibian_life_stage != 'EggMass') %>% 
  mutate(date = ymd(survey_date)) %>%
  mutate(year = year(date)) %>% 
  select(year, amphibian_species, 
         amphibian_life_stage, amphibian_number, lake_id, amphibian_location) %>%
  group_by(year, amphibian_life_stage) 
  
  
frogs_summary <- frogs_clean %>% 
  summarize(count = sum(amphibian_number, na.rm = TRUE))

```


```{r}
# Create a column graph of the above data frame, but do not incldue this graph in the knitted document 

years <- ggplot(data = frogs_summary, aes(x = year, y = count)) +
  geom_col(aes(fill = amphibian_life_stage), 
           width = 0.75, 
           position = "dodge",
           show.legend = TRUE) +
  scale_fill_manual(labels = c("Adult", "Sub-Adult", "Tadpole"),
                               values = c("yellowgreen",
                                          "seagreen3", "skyblue3")) +
  theme_bw() +
  labs(x = "Year", y = "Count", fill = 'Life Stage')

```

```{r}
# Create a second graph categorizing counts by the top five lakes 

top_lakes <- frogs_clean %>%
  filter(amphibian_life_stage != 'Tadpole') %>% 
  filter(amphibian_location == "in lake") %>% 
  group_by(lake_id) %>%
  summarize(count = sum(amphibian_number)) %>% 
  mutate(lake_id = as.character(lake_id)) %>%
  mutate(lake_id = paste("Lake", lake_id, sep = ' ')) %>% 
  mutate(lake_id = fct_reorder(lake_id, count)) %>% 
  slice_max(count, n = 5)

# Make the graph but do not show it in the knitted document

lakes <- ggplot(data = top_lakes, aes(x = reorder(lake_id, -count), 
                                      y = count)) +
  geom_col(fill = "darkseagreen4",
           width = 0.5,
           show.legend = FALSE) +
  theme_bw() +
  labs(x = "Lake ID", y = "Count")
```

```{r}
# Combine the two figures using cowplot::plot_grid

title <- ggdraw() + 
  draw_label(
    "Figure 1: Mountain Yellow-legged Frog Count by Lake and Life Stage",
    fontface = 'bold',
    x = 0,
    hjust = 0) +
  theme(plot.margin = margin(0, 0, 0, 7))


plots <- plot_grid(lakes + theme(axis.text.x = element_text(size = 7)),
                   years + theme(legend.position = "none",
                                 axis.title.y = element_blank()),
                   ncol = 2,
                   rel_widths = c(1.5, 1))

legend <- get_legend(years + 
                       theme(legend.position = "bottom"))


plot_grid(title, plots, legend, ncol = 1,
          rel_heights = c(0.1, 1, 0.1))

```


**Figure 1:** *The left graph provides a count of observed Mountain Yellow-legged Frogs (combining adults and sub-adults) across the five lakes (by Lake ID) with the highest recorded counts. The graph on the right provides a count of observed Mountain Yellow-legged Frogs across all lakes by life stage and year between 1995 and 2002.*

The visualization above shows that Mountain Yellow-legged Frog counts were relatively consistent across the five lakes with the highest observed numbers. It also indicates that the frogs are significantly spread out across all lakes. Overall, the observations of this speices increased over the study period with the hgihest numbers recorded in the final year of the study (2002). 

## **Data Citation** 

*Knapp, R.A., C. Pavelka, E.E. Hegeman, and T.C. Smith. 2020. The Sierra Lakes Inventory Project: Non-Native fish and community composition of lakes and ponds in the Sierra Nevada, California ver 2. Environmental Data Initiative. https://doi.org/10.6073/pasta/d835832d7fd00d9e4466e44eea87fab3*
