---
title: "Predicting Oxygen Concentration in California Seawater"
author: "Conner Smith"
date: "1/18/2022"
output: 
  html_document: 
    theme: flatly
---

```{r setup, include=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(AICcmodavg)
library(kableExtra)
library(equatiomatic)
```

## **Overview**

This analysis looks at a small subset of seawater sample data from CalCOFI and compares the performance of three competing linear regression models that predict oxygen saturation based on several physical and chemical variables, using AIC and cross validation.

The study was designed to explore the connection between O2 saturation of seawater off California’s coast and several physical and chemical variables. The 70+ year hydrographic time-series includes temperature, salinity, oxygen and phosphate observations. In 1961, nutrient analysis expanded to include silicate, nitrate and nitrite. 

## **Model Comparison**
```{r}
# Read in the data

sea <- read_csv(here("data", "calcofi_seawater_samples.csv"))
```
This analysis compares three models using different iterations of the variables recorded in the dataset. It compares model fit using AIC values and a 10-fold cross-validation test. 

##### **Table 1: Corrected AIC Values**
```{r}
# Create and compare two to three linear regressions

# Model 1 compares oxygen saturation as a function of water temperature, salinity, and phosphate concentration 

f1 <- o2sat ~ t_deg_c + salinity + po4u_m
mdl1 <- lm(f1, data = sea)

# Model 2 is the same but adds a term for depth  

f2 <- o2sat ~ t_deg_c + salinity + po4u_m + depth_m
mdl2 <- lm(f2, data = sea)

# Model 3 is exploratory and adds nitrite concentration while removing salinity

f3 <- o2sat ~ t_deg_c + po4u_m + no2u_m + depth_m
mdl3 <- lm(f3, data = sea)


# Create a table displaying all the corrected AIC values, the low3est value indicates the greatest model fit. 

aic_df <- data.frame() %>% 
  summarize(aic_mdl1 = AICc(mdl1),
            aic_mdl2 = AICc(mdl2),
            aic_mdl3 = AICc(mdl3)) 

aic_df %>% 
  kable(col.names = c("Model 1", "Model 2", "Model 3"), digits = 3) %>% 
  kable_styling(bootstrap_options = "striped", 
                position = "left", full_width = FALSE)

```

Model 2 has a lower corrected AIC value by more than two points, indicating it is a better fit than Model 1. This shows that depth is an important predictor variable for oxygen concentration. However, Model 3 has the lowest AIC value of all. Specifically, adding nitrite concentration and removing salinity improves the corrected AIC by 5 additional points. Model 3 will likely be the final model after the cross fold validation is performed.  

```{r}
# Perform 10-fold cross validation on the two models. 

folds <- 10
fold_vec <- rep(1:folds, length.out = nrow(sea))

# Using the same set.seed value from the lab.
set.seed(42)

sea_fold <- sea %>% 
  mutate(group = sample(fold_vec, size = n(), replace = FALSE))


# Set up the function to readily calculate RMSE 

calc_rmse <- function(x,y) {
  rmse_result <- (x - y)^2 %>% mean() %>% sqrt()
  return(rmse_result)
}

# The following chunk performs the 10-fold cross-validation for all three models. Removed the code from the lab where we just looked at the results of one fold. 

```

##### **Table 2: RMSE Values**
```{r}
# Calculate RMSE for all folds and then train the model on the entire data set. 
rmse_df <- data.frame()

for(i in 1:folds) {
  kfold_test_df <- sea_fold %>% 
    filter(group == i)
  kfold_train_df <- sea_fold %>% 
    filter(group != i)
  
  kfold_mdl1 <- lm(f1, data = kfold_train_df)
  kfold_mdl2 <- lm(f2, data = kfold_train_df)
  kfold_mdl3 <- lm(f3, data = kfold_train_df)
 
  kfold_pred_df <- kfold_test_df %>% 
    mutate(mdl1 = predict(kfold_mdl1, kfold_test_df),
           mdl2 = predict(kfold_mdl2, kfold_test_df),
           mdl3 = predict(kfold_mdl3, kfold_test_df))
  
  kfold_rmse <- kfold_pred_df %>% 
    summarize(rmse_mdl1 = calc_rmse(mdl1, o2sat),
              rmse_mdl2 = calc_rmse(mdl2, o2sat),
              rmse_mdl3 = calc_rmse(mdl3, o2sat))
  
  rmse_df <- bind_rows(rmse_df, kfold_rmse)
}

# Create a table to summarize the differences in average RMSE across all 10 folds for each of the three models. 

rmse_df %>% 
  summarize(rmse_mdl1 = mean(rmse_mdl1),
            rmse_mdl2 = mean(rmse_mdl2),
            rmse_mdl3 = mean(rmse_mdl3)) %>% 
  kable(col.names = c("Model 1", "Model 2", "Model 3"), digits = 3) %>% 
  kable_styling(bootstrap_options = "striped", 
                position = "left", full_width = FALSE)

```

This 10-fold cross-validation confirms that Model 3 has the best fit given the lowest RMSE value. This means that Model 3 predicts the actual observed values for oxygen concentration with the highest level of accuracy compared to the other models. The following shows the final equation generated from Model 3: 
```{r}
# Train the final model (model 3) on the entire data set (not the folds). 

final_mdl3 <- lm(f3, data = sea)

#Use `equatiomatic:: extract()` to generate an equation. 

extract_eq(model = final_mdl3, use_coefs = TRUE)
```

Here, the temperature (t_deg_c), depth (depth_m), and phosphate concentration (po4u_m) variables have an inverse relationship with oxygen saturation. Nitrite concentration (no2u_m) has a positive relationship with oxygen saturation. 

## **Data Citation**

*CalCOFI data are available for use without restriction. Data downloaded from https://calcofi.org/ccdata.html.*
