---
title: "Assignment 1 - Task 1"
author: "Conner Smith"
date: "1/13/2022"
output: 
  html_document: 
    theme: flatly
---

## **Overview**

This task explores amphibian abundance data recorded by the Sierra Lakes Inventory Project which ran from 1995-2002 and has supported research and management of Sierra Nevada aquatic ecosystems and their terrestrial interfaces.  

*Data citation: Knapp, R.A., C. Pavelka, E.E. Hegeman, and T.C. Smith. 2020. The Sierra Lakes Inventory Project: Non-Native fish and community composition of lakes and ponds in the Sierra Nevada, California ver 2. Environmental Data Initiative. https://doi.org/10.6073/pasta/d835832d7fd00d9e4466e44eea87fab3*

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(cowplot)
library(patchwork)
library(lubridate)

```

## **Analysis**

```{r}
# Read in xlsx file with 'read_xlsx

frogs <- readxl::read_xlsx(here("data", "sierra_amphibians.xlsx"))
```

... 

```{r}
# Filter for Mountain Yellow-legged Frogs (Species code: RAMU), excluding egg life stage (code: EggMass) and group by year and lifestage. 

frogs_clean <- frogs %>% 
  filter(amphibian_species == 'RAMU',
         amphibian_life_stage != 'EggMass') %>% 
  mutate(date = ymd(survey_date)) %>%
  mutate(year = year(date)) %>% 
  select(year, amphibian_species, 
         amphibian_life_stage, amphibian_number) %>%
  group_by(year, amphibian_life_stage) 
  
  
frogs_summary <- frogs_clean %>% 
  summarize(count = n())

```

### **Figure 1: Mountain Yellow-legged Frog Count By Year and Life Stage**

```{r}
# Create a column graph of the above data frame

ggplot(data = frogs_summary, aes(x = year, y = count)) +
  geom_col(aes(fill = amphibian_life_stage), 
           width = 0.75, 
           position = "dodge",
           show.legend = TRUE) +
  scale_fill_manual(labels = c("Adult", "Sub-Adult", "Tadpole"),
                               values = c("yellowgreen",
                                          "darkseagreen4", "skyblue3")) +
  theme_bw() +
  labs(x = "Year", y = "Count", fill = 'Life Stage')

```

#### **Figure 1:** *This graph provides a count of observed Mountain Yellow-legged Frogs by life stage and year between 1995 and 2002. There were no observations in 1998 or 1999. The highest number of frogs were observed in 2002.*
